+---
title: "2017_Set"
author: "M.Dubova"
date: '5 мая 2017 г '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Frequentist data analysis
## Marina Dubova

```{r load_libraries, echo=FALSE}
library(psych)
library(data.table)
library(foreign)
library(ggplot2)
library(lme4)
library(mlmRev)
library(lmerTest)
library(rstan)
library(ggmcmc)
library(bayesplot)
```

Data loading

```{r download_data}

setwd("/home/marina/Documents/Projects/Illusions_of_set/Experiment 1/data")
names <- c("IKA_Set_2017_05_22_1634.csv","NAT_Set_2017_05_05_1426.csv", "AOM_Set_2017_05_01_2019.csv", "BRO_Set_2017_05_05_1411.csv", "LDS_Set_2017_05_05_1327.csv", "DFX_Set_2017_04_15_1952.csv", "ASS_Set_2017_04_28_1416.csv", "ENE_Set_2017_04_28_1356.csv", "karina_Set_2017_04_28_1319.csv", "M.N_Set_2017_04_28_1457.csv", "OMK_Set_2017_04_15_2003.csv", "POL_Set_2017_04_15_1917.csv", "SAA_Set_2017_04_15_1800.csv", "SEV_Set_2017_04_28_1342.csv", "Simon_Set_2017_04_15_1935.csv", "Mil_Set_2017_04_29_1731.csv", "AAF_Set_2017_04_29_1410.csv", "GSA_Set_2017_05_17_1713.csv", "KLE_Set_2017_05_17_1645.csv", "AKS_Set_2017_05_17_1623.csv", "YGY_Set_2017_05_17_1609.csv", "ASV_Set_2017_05_17_1556.csv", "MB_Set_2017_05_17_1412.csv", "EZV_Set_2017_05_17_1327.csv", "ARM_Set_2017_05_16_2215.csv", "RBD_Set_2017_05_16_1316.csv")
data <- data.table()
for (i in 1:length(names)) {
  a <- read.csv(file = names[i])
  data <- rbind(data,a)
}
```

Making the data analyzable

1. Only test trials
2. Only relevant columns of PsychoPy output
3. Different subsets for Assimilative, Contrastive, and No illusion

```{r data_table, output=FALSE}
data_best <- data[,. (answer.keys, number_of_fixational_trials, size, trials_3.thisRepN, participant)]
#data_best$size <- as.factor(data_best$size)
#data_best$number_of_fixational_trials <- as.factor(data_best$number_of_fixational_trials)
data_best <- data_best[!(size=="NA")]
data_best <- data_best[(trials_3.thisRepN==0)]

data_right <- data_best[(answer.keys == "right")]
data_left <- data_best[(answer.keys == "left")]
data_left_right <- data_best[(answer.keys == "right"|answer.keys == "left")]
data_down <- data_best[(answer.keys == "down")]
levels(data_best$answer.keys) <- c("", "No illusion", "Contrastive illusion", "Assimilative illusion")
```

### Number of illusions

```{r plot0, echo = FALSE}
ggplot(data_best, aes(x = answer.keys)) + geom_bar() + ylab("Total amount of illusions") + xlab("") + ggtitle("Number of illusions")
```

### Different types of illusion and the size of difference between adaptive stimuli

```{r plot1, echo = FALSE}
ggplot(data_left_right, aes(x = size, fill = answer.keys)) + geom_bar()
ggplot(data_best, aes(x = size, fill = answer.keys)) + geom_bar(stat = "count", position = position_fill()) + ylab("Fraction of the type of illusions") + xlab("Size of difference between the setting stimuli (in individual differential thresholds)") + ggtitle("Fraction of illusions for different sizes of difference") + labs(fill = "Type of illusion")

ggplot(data_best, aes(x = answer.keys, fill = as.factor(number_of_fixational_trials))) + geom_bar(stat = "count", position = position_fill())
#ggplot(data_best, aes(x = answer.keys, fill = size)) + geom_bar(stat = "count", position = position_fill())
```

### Different types of illusion and the size of difference and number of adaptation trials

```{r plot2, echo = FALSE}
ggplot(data_left_right, aes(x = number_of_fixational_trials, fill = answer.keys)) + geom_bar(stat = "bin", position = position_fill())
```

### Contrastive illusion and the size of difference between adaptive stimuli

```{r plot3, echo = FALSE}
ggplot(data_left, aes(x = size)) + geom_bar() +  ylab("Total amount of illusions") + xlab("Size of difference between the setting stimuli (in individual differential thresholds)") + ggtitle("Number of contrastive illusions and difference in stimuli's sizes")
```

### Contrastive illusion and the number of adaptation trials

```{r plot4, echo = FALSE}
ggplot(data_left, aes(x = number_of_fixational_trials)) + geom_bar() + ylab("Total amount of illusions") + xlab("Number of fixational trials") + ggtitle("Number of contrastive illusions and number of fixational trials")
```

### Assimilative illusion and the size of difference between adaptive stimuli

```{r plot5, echo = FALSE}
ggplot(data_right, aes(x = size)) + geom_bar() + ylab("Total amount of illusions") + xlab("Size of difference between the setting stimuli (in individual differential thresholds)") + ggtitle("Number of assimilative illusions and difference in stimuli's sizes")
```

### Assimilative illusion and the number of adaptation trials

```{r plot6, echo = FALSE}
ggplot(data_right, aes(x = number_of_fixational_trials)) + geom_bar() + ylab("Total amount of illusions") + xlab("Number of fixational trials") + ggtitle("Number of assimilative illusions and number of fixational trials")
```

### Plots for 1-4 adaptation trials

```{r plot7, echo=FALSE}
data_right_num_1 <- data_right[(number_of_fixational_trials <= 4)]
data_right_num_2 <- data_right[(number_of_fixational_trials >= 5)]
ggplot(data_right_num_1, aes(x = size)) + geom_bar()
```

### Plots for 5-8 adaptation trials

```{r plot8, echo=FALSE}
ggplot(data_right_num_2, aes(x = size)) + geom_bar()
```

## Are the predictors correlated?

### Логистическая регрессия (для ассимилятивных: количество установочных проб ~ размер разницы)

```{r regression_1, echo = FALSE}

ggplot(data_right, aes(x = number_of_fixational_trials, y = size)) + geom_smooth(method = "glm", method.args = list(family = "gaussian"), aes(color = "gaussian"), se = TRUE)

```

### Логистическая регрессия (для контрастных: количество установочных проб ~ размер разницы)

```{r regression_2, echo = FALSE}

ggplot(data_left, aes(x = number_of_fixational_trials, y = size)) + geom_smooth(method = "glm", method.args = list(family = "gaussian"), aes(color = "gaussian"), se = TRUE)

```

### Логистическая регрессия (для обоих типов иллюзий: количество установочных проб ~ размер разницы)

``` {r regression_3, echo = FALSE}

ggplot(data_left_right, aes(x = number_of_fixational_trials, y = size, col = answer.keys)) + geom_smooth(method = "glm", method.args = list(family = "gaussian"), se = TRUE)

```

### Data with absolute value of difference between stimuli

```{r new_data, echo = FALSE}
data_best_2 <- data[,. (answer.keys, number_of_fixational_trials, size, trials_3.thisRepN, participant)]
data_best_2 <- data_best_2[!(size=="NA")]
data_best_2 <- data_best_2[(trials_3.thisRepN==0)]
levels(data_best_2$answer.keys) <- c("", "No illusion", "Contrastive illusion", "Assimilative illusion")


c1 <- aggregate(data$rating_up.response, by=list(Category=data$participant), FUN=mean, na.rm = TRUE) 
c2 <- aggregate(data$rating_down.response, by=list(Category=data$participant), FUN=mean, na.rm = TRUE)
c1$sum <- c1$x + c2$x
c1$mean <- (c1$sum)/2
c1 <- c1[rep(1:nrow(c1), each = 24),]
data_best_2$size <- data_best$size * c1$mean

data_right_2 <- data_best_2[(answer.keys == "Assimilative illusion")]
data_left_2 <- data_best_2[(answer.keys == "Contrastive illusion")]
data_left_right_2 <- data_best_2[(answer.keys == "Assimilative illusion"|answer.keys == "Contrastive illusion")]
data_down_2 <- data_best_2[(answer.keys == "No illusion")]

data_best_2$size_in_thresholds <- data_best$size
data_best_2$answer <- data_best_2$answer.keys
levels(data_best_2$answer.keys) <- c("NA", "0", "-1", "1")
data_best_2$answer.keys <- as.numeric(data_best_2$answer.keys)
write.csv(data_best_2, file = "home/marina/Documents/Projects/Illusions_of_set/Set 1/Experiment/data/MyData.csv")
```

### GEOM POINT

``` {r plot9, echo = FALSE}
ggplot(data_left_right_2, aes(x = size, y = number_of_fixational_trials, col = answer.keys)) + geom_jitter()
ggplot(data_left_2, aes(x = size)) + geom_histogram(binwidth = 0.1) + xlab("Absolute value of differece between setting stimuli (in cm)") + ylab("Total amount of illusions") + ggtitle("Number of contrastive illusions and absolute value of difference")

# DENSITIES
ggplot(data_best_2, aes(x = size, fill = answer.keys)) + geom_density(alpha=0.4) + xlab("Absolute value of differece between setting stimuli (in cm)") + ylab("Estimated density") + ggtitle("Estimated density of distributions") + labs(fill= "Type of illusion")

# DISTRIBUTIONS
ggplot(data_best_2, aes(x = size, fill = answer.keys)) + geom_histogram(binwidth = 0.1) + xlab("Absolute value of differece between setting stimuli (in cm)") + ylab("Total amount of illusions") + ggtitle("Number of illusions and absolute value of difference") + facet_wrap(~ answer.keys) + labs(fill= "Type of illusion")


ggplot(data_left_right_2, aes(x = size, fill  = answer.keys)) + geom_histogram(alpha = 0.3, binwidth = 0.1, aes(y = ..density..), position = 'identity') + xlab("Absolute value of differece between setting stimuli (in cm)") + ylab("Total amount of illusions") + ggtitle("Number of assimilative illusions and absolute value of difference")


ggplot(data_right_2, aes(x = size)) + geom_histogram(binwidth = 0.1) + xlab("Absolute value of differece between setting stimuli (in cm)") + ylab("Total amount of illusions") + ggtitle("Number of assimilative illusions and absolute value of difference")

###ggplot(data_left_right_2, aes(x=size, fill = answer.keys)) + geom_histogram(aes(y=..count../sum(..count..)))

ggplot(data = data_left_right_2,aes(x=size, fill = answer.keys)) + geom_density(alpha = 0.3) + geom_dotplot(dotsize = 0.5,aes(size=2, fill=answer.keys))
#+ labs(fill="Type of illusion", linetype="Type of illusion")
``` 

## Mixed logistic regression

### Модель для дифференциальных порогов

``` {r lrem2}

model1 <- glmer(answer.keys ~ size + number_of_fixational_trials + (1|participant), data = data_best, family = binomial) # doesn't converge with individual effects of num_of_trials
summary(model1)
model10 <- glmer(answer.keys ~ 1 + (1|participant), family = binomial, data = data_best)
summary(model10)
anova(model10, model1)
```

### Модель со смешанными эффектами для абсолютной разницы

``` {r lrem1}
model2 <- glmer(answer.keys ~ size + number_of_fixational_trials + (1|participant) + (size|participant), data = data_left_right_2, family = binomial)
summary(model1)
model20 <- glmer(answer.keys ~ 1 + (1|participant), data = data_left_right_2, family = binomial)
summary(model0)
anova(model2, model20)
```

